Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-04 architecture.tex
done or the day
work in progress
cleaning up stuff

%!TEX root=restructured.tex

@@

>>>>

%!TEX root=paper.tex

    \begin{figure*}[ht!]
      \centering
      \includegraphics[width=0.8\linewidth]{interceptor}
      \caption{The first thing that needs to be done, is to decorate the application object with an INTERCEPTOR}
      \label{fig:sep}
    \end{figure*}

@@

The fact that the \tool itself is being developed in Python using Flask makes binding to the services of a to-be monitored application developed with the same technologies easy and intuitive. However, by writing different backends that do the measurement and monitoring, the frontend can be reused.

### 2018-04-04 discussion.tex
work in progress
cleaning up stuff

-   The fact that the \tool itself is being developed in Python using Flask makes binding to the services of a to-be monitored application developed with the same technologies easy and intuitive. However, by writing different backends that do the measurement and monitoring, the frontend can be reused.

@@

>>>>

@@

- Although predicting the performance of the application would require advanced statistical work, it can be succesfully used as an advanced warning system for the overall performance trend for given endpoints.

### 2018-04-04 introductory.tex
work in progress
cleaning up stuff

\begin{abstract}
\todo{Update abstract \& title}
  Tens of thousands of web applications are written in Flask, a Python-based web framework. Despite a rich ecosystem of extensions, there is none that supports the developer in gaining insight into the evolving performance of their service. In this paper, we introduce \tool, a library that addresses this problem. We present the ease with which the library can be integrated in an already existing web application, discuss some of the visualization perspectives that the library provides and point to some future challenges for similar libraries.

@@

>>>>

\begin{abstract}
\todo{Update abstract \& title}
  Python is one of the fastest growing programming languages of the moment. Flask, a Python-based web framework is the technology chosen by tens of thousands of web applications. Many of these Flask-based projects are small websites which can not afford to deploy an advanced service monitoring solution because they are too complicated to install.

  In this paper, investigate how simple can it be to start monitoring a Flask API by developing and presenting \tool, a library that aims to provide insight into the utilization and performance of evolving services. We present the ease with which the library can be integrated in an already existing web application, discuss some of the visualization perspectives that the library provides and point to some future challenges for similar libraries.

@@

  \ml{add more numbers that show how impressive is our API}

### 2018-04-04 overhead.tex
cleaning up stuff

\documentclass[conference]{IEEEtran}
	\pdfpagewidth=8.5truein
	\pdfpageheight=11truein

\usepackage{graphicx}
\usepackage{subfig}
\usepackage[bookmarks=false]{hyperref}
\usepackage{xspace}
\usepackage{listings}
\usepackage[usenames, dvipsnames]{color}
\usepackage{amssymb}

\usepackage{booktabs}

\graphicspath{{./img/}}

%\hyphenation{op-tical net-works semi-conduc-tor}

\input{useful_defines}

\begin{document}

%
\title{Agile Monitoring of Evolving Web Services}

\author{
\IEEEauthorblockN{The Author, The Other Author}\\
Institute\\
University, Country\\
Email: \{author\}@university.domain 
}

% make the title area
\maketitle

\input{introductory}

\input{loc1}

\input{utilization}

\input{performance}

\input{grouping}

\input{evolution}

\input{testing}

\input{architecture}

\section{Performance Perspectives}
\label{sec:views}

@@

  %In the remainder of the paper we present several of these perspectives.

  % The second endpoint conists of two parts, one of them being a table that shows for every monitored endpoint the number of hits it has gotten, the time it was last accessed and its average execution time. The second part is a view with four graphs which show:
  % \begin{itemize} 
  %   \item A heatmap of the total number of requests to the monitored endpoints
  %   \item A stacked bar chart that shows the total number of requests to the monitored endpoints per endpoint per day
  %   \item A boxplot graph showing the average execution time per version of the web service
  %   \item A boxplot graph showing the average execution time for every monitored endpoint
  % \end{itemize}

\todo{evaluate which figures to keep and ideally take new screenshots for the ones we keep; consider adding two more views (like Fig. 3.2 of Patrick's thesis)}

% In the Zeeguu case study, one of the slowest endpoints, and one with the highest variability as shown in \Fref{fig:ep} is \epFeedItems: it retrieves a list of recommended articles for a given user. However, since a user can be subscribed to anything from one to three dozen article sources, and since the computation of the difficulty is personalized and it is slow, the variability in time among users is likely to be very large.

% \niceseparator

%\section{Preemptive Monitoring}
%\label{sec:testing}
%
%%\todo{discuss here idea of using unit testing as an early performance indicator for evolving services (`preemptive monitoring'?); add a side-by-side figure comparing unit testing vs live system performance per version, maybe calculate error in trend curves? find a better term than prediction for the section title (plus introduction)}
%
%\todo{Similar to the idea of augmenting service monitoring with online testing~\cite{metzger2010proactive}, i.e.~testing service-based applications by using dedicated test input in parallel to its normal use and operation. The difference is that we take advantage of the capability of Travis to create an emulated live environment for integration testing purposes, and use unit testing as the dedicated test input in order to measure performance. What we are going provide evidence for in the following is that we can indeed use this preemptive monitoring concept as an early performance indicator for evolving services.}
%
%The developer needs to define the unit test folder of the project, the URL that Travis should use for reporting its results, and the number of iterations for each unit test should be executed in order to generate

\section{Evaluation}
\label{sec:evaluation}

%\todo{evaluation of the \tool across the following dimensions: ease of use, effectiveness, overhead to the system; for the last part we need to do the overhead measurement experiment, for the other two we base it on the case study and bring over whatever is necessary from the discussion section}

\input{discussion}

\input{related-work}

\section{Conclusion and Future Work}
\label{sec:conclusions}

\todo{update as appropriate}

In this paper we have shown that it is possible to create a monitoring solution which provides basic insight into web service utilization and performance  with very little effort from the developer. The user group that we are aiming for with this work is application developers using Flask and Python to build web applications with limited or no budget for implementing their own monitoring solutions. The emphasis is in allowing such users to gain insight into how the performance of the service evolves together with the application itself. We believe that the same architecture, and lessons can be applied to other frameworks and other languages.

In the future, we plan to perform case studies with other sytstems, with the goal of discover other needs and to wean out the less useful visualizations in the \tool. We plan to also extend the tool towards supporting multiple deployments of the same applications across multiple nodes (e.g. for the situations where the application is deployed together with a load balancer). Finally, we plan to integrate \tool with unit testing as a complementary source of information about performance evolution.

\newpage
\appendix

  \subsection{Changing the Default /dashboard Endpoint }

  THIS STUFF SHOULD BE MOVED TO THE APPENDIX. And in a footnote we just mention that ... \ins{REQUIREMENT}: The system must provide an easy way to add more specific configurations if needed. A way of overwriting the common sense defaults.

  Further configuration is not needed possible by adding additional statements before the binding definition, for example,

  \begin{lstlisting}[style=custompython]
  ...
  dashboard.config.link = 'mydashboard'
  dashboard.bind(flask_app)
  \end{lstlisting}

  allows for a custom route (\code{/mydashboard}) to the dashboard to be defined by the programmer. An external configuration file can be used instead, or in addition to this:

  \begin{lstlisting}[style=custompython]
  ...
  dashboard.config.from_file('config.cfg')
  ...
  \end{lstlisting}

  \subsection{For those who don't use GIT}

    \textit{Manual} version control requires the developer to tag each new version of the application with an appropriate version identifier~\cite{papazoglou2011managing} using the \code{APP\_VERSION} configuration parameter, for example by adding to the configuration file:

      \begin{lstlisting}[style=custompython]
      [dashboard]
      APP_VERSION=<versionID>
      \end{lstlisting}

% references section

\bibliographystyle{abbrv}
\bibliography{paper}

% that's all folks
\end{document}

>>>>

%!TEX root=paper.tex

\section{Overhead of the \tool}
\label{sec:overhead}

@@

### 2018-04-04 paper.tex
cleaned up the performance testing part
cleaning up stuff

%
\title{Agile Monitoring of Evolving Web Services}

@@

@@

\section{Performance Perspectives}
\label{sec:views}

\ins{PERFORMANCE DISCUSSION: this introspection, which looks up endpoints, happens only once at the startup of the service, so it is going to affect the API startup performance, but not the actual endpoint performance. In our case, the overhead here was quite small: TO MEASURE}.

  %In the remainder of the paper we present several of these perspectives.

  % The second endpoint conists of two parts, one of them being a table that shows for every monitored endpoint the number of hits it has gotten, the time it was last accessed and its average execution time. The second part is a view with four graphs which show:
  % \begin{itemize} 
  %   \item A heatmap of the total number of requests to the monitored endpoints
  %   \item A stacked bar chart that shows the total number of requests to the monitored endpoints per endpoint per day
  %   \item A boxplot graph showing the average execution time per version of the web service
  %   \item A boxplot graph showing the average execution time for every monitored endpoint
  % \end{itemize}

\todo{evaluate which figures to keep and ideally take new screenshots for the ones we keep; consider adding two more views (like Fig. 3.2 of Patrick's thesis)}

% In the Zeeguu case study, one of the slowest endpoints, and one with the highest variability as shown in \Fref{fig:ep} is \epFeedItems: it retrieves a list of recommended articles for a given user. However, since a user can be subscribed to anything from one to three dozen article sources, and since the computation of the difficulty is personalized and it is slow, the variability in time among users is likely to be very large.

% \niceseparator

%\section{Preemptive Monitoring}
%\label{sec:testing}
%
%%\todo{discuss here idea of using unit testing as an early performance indicator for evolving services (`preemptive monitoring'?); add a side-by-side figure comparing unit testing vs live system performance per version, maybe calculate error in trend curves? find a better term than prediction for the section title (plus introduction)}
%
%\todo{Similar to the idea of augmenting service monitoring with online testing~\cite{metzger2010proactive}, i.e.~testing service-based applications by using dedicated test input in parallel to its normal use and operation. The difference is that we take advantage of the capability of Travis to create an emulated live environment for integration testing purposes, and use unit testing as the dedicated test input in order to measure performance. What we are going provide evidence for in the following is that we can indeed use this preemptive monitoring concept as an early performance indicator for evolving services.}
%
%The developer needs to define the unit test folder of the project, the URL that Travis should use for reporting its results, and the number of iterations for each unit test should be executed in order to generate

\section{Evaluation}
\label{sec:evaluation}

%\todo{evaluation of the \tool across the following dimensions: ease of use, effectiveness, overhead to the system; for the last part we need to do the overhead measurement experiment, for the other two we base it on the case study and bring over whatever is necessary from the discussion section}

>>>>

%
\title{Agile Monitoring of Evolving Web Services: For the Little Man}

@@

\input{overhead}

@@
